`r params$project_title`

### 2021-11-18 reports/rmarkdown/01_data_pre_processing.Rmd
update

```{r, echo = FALSE}
# Load metadata
dataset_names <- c("ASV_tissue_B2.csv",   # Tissue, Boston run 2 (95 samples)
                     "ASV_tissue_B1.csv", # Tissue, Boston run 1 (1 sample)
                     "ASV_CVL_V2_B1.csv", # CVL V2, Boston run 1 (27 samples)
                     "ASV_CVL_V2_B2.csv", # CVL V2, Boston run 2 (49 samples)
                     "ASV_CVL_V2_C.csv",  # CVL V2, CTMR         (62 samples)
                     "ASV_CVL_V3.csv")

@@

```{r, echo = FALSE}
temp <- map(datasets_, ~.x  %>%
            select(Sequence, Seq_ID, Kingdom:Species, everything()) %>%
            filter(!(is.na(.$Kingdom))) %>% # Removes any kingdom being NA
            other_taxa.fun(.) %>% # create "other" taxonomy 
            genus_lacto.fun(.)  
            ) %>% # create genus lacto taxonomy 
        {. ->> temp_ } %>% 
        map(., ~aggregate.fun(., "Genus_lacto"))

# Extract taxonomic information to file
seq_taxa <- bind_rows(temp_) %>% select(1:11) %>% unique() 
#write.csv(seq_taxa, paste0(PATH,"results/","SeqID_to_gen-lacto_tax-oth",".csv"),row.names = F )

@@

  taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
  taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
  taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
  taxa[!not_complete] <- sps[!not_complete]
  taxa <- sub("_"," ",taxa)
  use <- !is.na(taxa) & (taxa != "NA")
  temp <- as.matrix( x[,-c(1,(ncol(x)-7):ncol(x))] )
  temp <- rowsum(temp[use,],taxa[use])
  colnames(temp) <- colnames(x)[-c(1,(ncol(x)-7):ncol(x))]
  return(temp)
})
names(datasets) <- dataset_names
lapply(datasets,dim)
lapply(datasets,colnames)
#arrange all datasets to have the same rownames with all bacteria
common_microbes <- unique(unlist(lapply(datasets,rownames)))
datasets <- lapply(datasets, function(x) {
  temp <- matrix(0,nrow = length(common_microbes),
                 ncol = ncol(x),
                 dimnames = list(common_microbes,colnames(x)))
  temp[rownames(x),colnames(x)] <- x
  return(temp)
})
lapply(datasets,dim)
# datasets <- datasets[3] #in case you wanna test in 1 dataset
# Transform phyloseq data into normalized matricies ( log2[CP1K] )
# datasets <- lapply(datasets,function(x){
#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
# })
# lapply(datasets,dim)
# Extract taxonomic information to file
taxa_info <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"data/",x))
  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
  temp[is.na(temp)] <- "NA"
  return(t(temp))
})
taxa_info <- t(as.data.frame(taxa_info))
rownames(taxa_info) <- sub("other.","other ",rownames(taxa_info))
rownames(taxa_info) <- sub("[.][0123456789]","other ",rownames(taxa_info))
sps <- apply( taxa_info[,6:7] ,1,function(x) paste(x,collapse ="_"))
not_complete <- grepl("_NA",sps)
taxa <- apply( taxa_info ,1,function(x) paste(x,collapse ="_"))
taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
taxa[!not_complete] <- sps[!not_complete]
taxa <- sub("_"," ",taxa)
use <- !is.na(taxa) & (taxa != "NA")
taxa[taxa == "NA"] <- "other"
rownames(taxa_info) <- taxa
taxa_info <- taxa_info[! duplicated(rownames(taxa_info)),]
write.csv2(taxa_info, paste0(PATH,"results/taxonomy.csv"),row.names = T)
```

@@

datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3.csv"]]) / colSums(datasets[["ASV_CVL_V3.csv"]]) * 1000 + 1 )),4)
datasets[["ASV_tissue_V3"]] <- round(t( log2( t(datasets[["ASV_tissue_V3"]]) / colSums(datasets[["ASV_tissue_V3"]]) * 1000 + 1 )),4)
datasets[["ASV_CVL_V2_raw"]] <- round(t( log2( t(datasets[["ASV_CVL_V2_raw"]]) / colSums(datasets[["ASV_CVL_V2_raw"]]) * 1000 + 1 )),4)
datasets[["ASV_CVL_V2"]] <- round( datasets[["ASV_CVL_V2"]] ,4)

>>>>

```{r, echo = FALSE}
# Load metadata
dataset_names <- c("ASV_tissue_V3_B2.csv",   # Tissue, Boston run 2 (95 samples)
                     "ASV_tissue_V3_B1.csv", # Tissue, Boston run 1 (1 sample)
                     "ASV_CVL_V2_B1.csv", # CVL V2, Boston run 1 (27 samples)
                     "ASV_CVL_V2_B2.csv", # CVL V2, Boston run 2 (49 samples)
                     "ASV_CVL_V2_S.csv",  # CVL V2, CTMR         (62 samples)
                     "ASV_CVL_V3_B1.csv") # CVL V3, Boston run 1 (111 samples)

@@

```{r, echo = FALSE}
temp <- map(datasets_, ~.x  %>%
            select(Sequence, Seq_ID, Kingdom:Species, everything()) %>%
            filter(!(is.na(.$Kingdom))) %>% # Removes any kingdom being NA
            other_taxa.fun(.) %>% # create "other" taxonomy 
            genus_lacto.fun(.)  # create genus lacto taxonomy
            ) %>%  
        {. ->> temp_ } %>% 
        map(., ~aggregate.fun(., "Genus_lacto")) # agglomerate on genus/lacto spp. lvl.

# Extract taxonomic information to file
seq_taxa <- bind_rows(temp_, .id = "dataset") %>% select(1:12) %>% unique() 
#write.csv(seq_taxa, paste0(PATH,"results/","SeqID_to_gen-lacto_tax-oth",".csv"),row.names = F )

# make long format raw non agglom ASV files
# remove_from_CVL_V2 <- intersect(colnames(temp_[[5]]), colnames(temp_[[3]]))[-c(1:11)]
# names <- c("ASV_tissue_V3_Boston run2","ASV_tissue_V3_Boston run1", "ASV_CVL_V2_Boston run1", 
#                      "ASV_CVL_V2_Boston run2", "ASV_CVL_V2_Stockholm", "ASV_CVL_V3_Boston run 1") 
# ASV_long <- temp_ %>%
#   modify_at(., c("ASV_CVL_V2_C.csv"), ~select(.x, -any_of(remove_from_CVL_V2))) %>% 
#   map(., ~pivot_longer(.x, matches("^P\\d"), names_to = "ID", values_to = "Reads")) %>%
#   set_names(names ) %>%
#   bind_rows(., .id = "dataset") %>%
#   separate(., dataset, into=c(NA, "Sample_type", "Visit","Place_sequenced"), sep = "_", remove = F) %>% 
#   mutate(dataset = str_extract(.$dataset, "CVL_V\\d|tissue_V3")) %>% 
#   group_split(dataset, .keep=F) %>% set_names(c("tissue", "CVL_V2", "CVL_V3"))
# 
# saveRDS(ASV_long, paste0(PATH,"results/","ASV_all_data_long_format.RDS"))

@@

  taxa <- apply( temp ,1,function(x) paste(x,collapse ="_"))
  taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
  taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
  taxa[!not_complete] <- sps[!not_complete]
  taxa <- sub("_"," ",taxa)
  use <- !is.na(taxa) & (taxa != "NA")
  temp <- as.matrix( x[,-c(1,(ncol(x)-7):ncol(x))] )
  temp <- rowsum(temp[use,],taxa[use])
  colnames(temp) <- colnames(x)[-c(1,(ncol(x)-7):ncol(x))]
  return(temp)
})
names(datasets) <- dataset_names
lapply(datasets,dim)
lapply(datasets,colnames)

#arrange all datasets to have the same rownames with all bacteria
common_microbes <- unique(unlist(lapply(datasets,rownames)))
datasets <- lapply(datasets, function(x) {
  temp <- matrix(0,nrow = length(common_microbes),
                 ncol = ncol(x),
                 dimnames = list(common_microbes,colnames(x)))
  temp[rownames(x),colnames(x)] <- x
  return(temp)
})
lapply(datasets,dim)

# datasets <- datasets[3] #in case you wanna test in 1 dataset
# Transform phyloseq data into normalized matricies ( log2[CP1K] )
# datasets <- lapply(datasets,function(x){
#     x <- t(log2( t(x) / colSums(x) * 1000 + 1))
# })
# lapply(datasets,dim)

# Extract taxonomic information to file
taxa_info <- lapply(dataset_names, function(x){
  x <- read.csv( paste0(PATH,"data/",x))
  temp <- as.matrix(x[,c((ncol(x)-6):ncol(x))])
  temp[is.na(temp)] <- "NA"
  return(t(temp))
})
taxa_info <- t(as.data.frame(taxa_info))
rownames(taxa_info) <- sub("other.","other ",rownames(taxa_info))
rownames(taxa_info) <- sub("[.][0123456789]","other ",rownames(taxa_info))
sps <- apply( taxa_info[,6:7] ,1,function(x) paste(x,collapse ="_"))
not_complete <- grepl("_NA",sps)
taxa <- apply( taxa_info ,1,function(x) paste(x,collapse ="_"))
taxa[not_complete] <- sub("_NA.*","",taxa[not_complete])
taxa[not_complete] <- sub(".*_","other ",taxa[not_complete])
taxa[!not_complete] <- sps[!not_complete]
taxa <- sub("_"," ",taxa)
use <- !is.na(taxa) & (taxa != "NA")
taxa[taxa == "NA"] <- "other"
rownames(taxa_info) <- taxa
taxa_info <- taxa_info[! duplicated(rownames(taxa_info)),]
write.csv2(taxa_info, paste0(PATH,"results/taxonomy.csv"),row.names = T)

@@

# normalized files
datasets[["ASV_CVL_V3.csv"]] <- round( t(log2( t(datasets[["ASV_CVL_V3.csv"]]) / colSums(datasets[["ASV_CVL_V3.csv"]]) * 1000 + 1 )),4)
datasets[["ASV_tissue_V3"]] <- round(t( log2( t(datasets[["ASV_tissue_V3"]]) / colSums(datasets[["ASV_tissue_V3"]]) * 1000 + 1 )),4)
datasets[["ASV_CVL_V2_raw"]] <- round(t( log2( t(datasets[["ASV_CVL_V2_raw"]]) / colSums(datasets[["ASV_CVL_V2_raw"]]) * 1000 + 1 )),4)
datasets[["ASV_CVL_V2"]] <- round( datasets[["ASV_CVL_V2"]] ,4)
